The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-07-22 script/de/chapters/digitalsignatures.tex
script SHA-1 update

Über mehrere Jahre gab es keine neuen Attacken gegen Hashverfahren,
und allgemein wurde den Kandidaten, die in der Praxis bislang keine
Schwächen in ihrer Struktur gezeigt hatten 
(zum Beispiel \index{SHA-1} SHA-1\footnote{%
  SHA-1 \index{SHA-1} ist eine in den Standards FIPS 180-1 (durch die
  US-Behörde NIST), ANSI X9.30 Part 2 und
  \cite{FIPS186} spezifizierte 160-Bit Hashfunktion.\\
  SHA bedeutet ``Secure Hash Algorithm'' und wird häufig benutzt, z.B. 
  mit DSA, RSA oder ECDSA.\\
  Der aktuelle Standard \cite{FIPS180-2} definiert vier sichere Hashverfahren
  -- SHA-1, SHA-256, SHA-384 und SHA-512.
  Für diese Hashalgorithmen sind in der Testsuite FIPS 140-2 auch
  Validierungstests definiert.

@@

Die Veröffentlichung von Details zu den Resultaten der
chinesischen Kryptologen wurde bisher nur angekündet: Nach ihren 
Angaben kann man
Kollisionen von SHA-1 mit einem Arbeitsaufwand von $2^{69}$ finden.
Damit wäre gezeigt, dass SHA-1 kryptographische Schwächen besitzt, 
da die Konstruktion von SHA-1 an sich einen Mindestarbeitsaufwand 
für die Kollisionssuche von etwa $2^{80}$ gewährleisten sollte.
Der Wert $2^{69}$ ist dabei eine auf theoretischen Hochrechnungen 
beruhende Vorhersage. 
Die Sicherheit bereits erstellter Signaturen wird durch den geschilderten
Kollisionsangriff aber nicht gefährdet.

@@

Das NIST hat schon vor Bekanntwerden der neuen Ergebnisse angekündigt,
SHA-1 in den näch"-sten Jahren auslaufen zu lassen. Es ist daher zu
empfehlen, für neue Produkte zur Erstellung von Sig"-naturen SHA-1 nicht
mehr zu verwenden.

@@

\bibitem[FIPS180-2]{FIPS180-2} U.S. Department of Commerce/N.I.S.T. ,
    \index{FIPS180-2} \\
    {\em Secure Hash Standard (SHS)}, \\
    August 1, 2002.\\
    (FIPS 180-2 will supersede FIPS 180-1 beginning February 1, 2003.)

@@

@@

>>>>

Über mehrere Jahre gab es keine neuen Attacken gegen Hashverfahren,
und allgemein wurde den Kandidaten, die in der Praxis bislang keine
Schwächen in ihrer Struktur gezeigt hatten 
(zum Beispiel \index{SHA-1} SHA-1\footnote{%
  SHA-1 \index{SHA-1} ist eine in den Standards FIPS 180-1 (durch die
  US-Behörde NIST), ANSI X9.30 Part 2 und
  \cite{FIPS186} spezifizierte 160-Bit Hashfunktion.\\
  SHA bedeutet ``Secure Hash Algorithm'' und wird häufig benutzt, z.B. 
  mit DSA, RSA oder ECDSA.\\
  Der aktuelle Standard \cite{FIPS180-3} definiert vier sichere Hashverfahren
  -- SHA-1, SHA-256, SHA-384 und SHA-512.
  Für diese Hashalgorithmen sind in der Testsuite FIPS 140-2 auch
  Validierungstests definiert.

@@

Die zunächst veröffentlichten Resultate reduzierten den erwarteten Aufwand für
die Suche nach einer SHA-1 Kollision von $2^{80}$ (brute force) auf $2^{69}$
\cite{Wang2005}.  In der Folge wurden Verfahren angekündigt, die den Aufwand
weiter auf $2^{63}$ \cite{Wang2005b} und $2^{52}$ \cite{McDonald2009} reduzieren
sollen.  Damit wäre der Kollisionsangriff in den Bereich des praktisch möglichen
gerückt, denn ähnliche Aufwände wurde in der Vergangenheit schon realisiert (s.\
\ref{Brute-force-gegen-Symmetr}).

Die Sicherheit bereits erstellter Signaturen wird durch den geschilderten
Kollisionsangriff aber nicht gefährdet.

@@

Das U.S. National Institute of Standards and Technology (NIST)\index{NIST} hat
schon vor Bekanntwerden der neuen Ergebnisse angekündigt, SHA-1 in den
näch"-sten Jahren auslaufen zu lassen. Es ist daher zu empfehlen, für neue
Produkte zur Erstellung von Sig"-naturen SHA-1 nicht mehr zu verwenden. Die
SHA-2 Familie \cite{FIPS180-3} bietet stärkere Verfahren.  Um den neuen
Erkenntnisse in der Kryptoanalyse von Hashfunktionen Rechnung zu tragen hat das
NIST 2008 einen Wettbewerb gestartet, in dem eine neue Hashfunktion ``SHA-3''
ausgewählt werden
soll.\footnote{\url{http://csrc.nist.gov/groups/ST/hash/sha-3/}} Der Wettbewerb
soll im Jahr 2012 enden.

@@

\bibitem[FIPS180-3]{FIPS180-3} U.S. Department of Commerce/N.I.S.T. ,
    \index{FIPS180-3} \\
    {\em Secure Hash Standard (SHS)}, \\
    October 2008.\\
    (FIPS 180-3 supersedes FIPS 180-2.)

@@

\bibitem[McDonald2009]{McDonald2009} Cameron McDonald, Philip Hawkes, Josef Pieprzyk, 
    \index{McDonald 2009} \\
    {\em Differential Path for SHA-1 with complexity $O(2^{52})$}, \\
    \url{http://eprint.iacr.org/2009/259}

@@

\bibitem[Wang2005]{Wang2005} Xiaoyun Wang, Yiqun Yin, Hongbo Yu, 
    \index{Wang 2005} \\
    {\em Finding Collisions in the Full SHA-1}, \\
    Advances in Cryptology-Crypto 2005, LNCS 3621: 17-36, 2005.

\bibitem[Wang2005b]{Wang2005b}  Xiaoyun Wang, Andrew Yao and Frances Yao, 
    \index{Wang 2005} \\
    {\em New Collision Search for SHA-1}, \\
    Crypto 2005 Rump Session \\
    \url{http://www.iacr.org/conferences/crypto2005/rumpSchedule.html}

### 2009-07-22 script/en/chapters/digitalsignatures.tex
script SHA-1 update

During the past several years no new attacks against hash algorithms
came up, and so the candidates that had not yet shown any weaknesses
in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
  SHA-1 \index{SHA-1} is a 160 bit hash function specified in FIPS 180-1
  (by NIST), ANSI X9.30 Part 2 and \cite{FIPS186}.\\
  SHA means Secure Hash Algorithm, and is widely used, 
  e.g. with DSA, RSA or ECDSA.\\
  The current standard \cite{FIPS180-2} defines four secure hash algorithms
  -- SHA-1, SHA-256, SHA-384, and SHA-512.
  For these hash algorithms there are also validation tests defined in the
  test suite FIPS 140-2.

@@

Right now the details of the results of the Chinese cryptographers are
only announced, but not revealed completely: They stated that collision
attacks on SHA-1 can be found with a work load of $2^{69}$.
This would mean that SHA-1 has cryptographic weaknesses, 
because the design of SHA-1 should ensure a work load for collision search
of about $2^{80}$.
The value of $2^{69}$ is currently a prognosis based on theoretic forecasts. 
The security of already created digital signatures is not affected by the
described attacks.

@@

Already before Crypto 2004 NIST announced, to discontinue SHA-1 
in the next few years. So it is recommended not to use SHA-1 for
new products generating digital signatures.

>>>>

During the past several years no new attacks against hash algorithms
came up, and so the candidates that had not yet shown any weaknesses
in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
  SHA-1 \index{SHA-1} is a 160 bit hash function specified in FIPS 180-1
  (by NIST), ANSI X9.30 Part 2 and \cite{FIPS186}.\\
  SHA means Secure Hash Algorithm, and is widely used, 
  e.g. with DSA, RSA or ECDSA.\\
  The current standard \cite{FIPS180-3} defines four secure hash algorithms
  -- SHA-1, SHA-256, SHA-384, and SHA-512.
  For these hash algorithms there are also validation tests defined in the
  test suite FIPS 140-2.

@@

The initially published result reduce the expected complexity for one SHA-1
collision search from $2^{80}$ (brute force) to $2^{69}$
\cite{Wang2005}. More recent announcements claim to futher reduce the required
effort to $2^{63}$ \cite{Wang2005b} and $2^{52}$ \cite{McDonald2009}. This would
bring collision attacks into the practial real, as similar efforts have been
mastered in the past (s.\ \ref{Brute-force-gegen-Symmetr}).

@@

Already before Crypto 2004 the U.S. National Institute of Standards and
Technology (NIST)\index{NIST} announced, to discontinue SHA-1 in the next few
years. So it is recommended not to use SHA-1 for new products generating digital
signatures. The SHA-2 family \cite{FIPS180-3} provides stronger algorithms. To
address new findings in crytanalysis, NIST has opened a competition to develop a
new cryptographic hash algorithm ``SHA-3'' in 2008. The competition is expected
to end in 2012.\footnote{\url{http://csrc.nist.gov/groups/ST/hash/sha-3/}}
